Evidence for an Intrinsic Color Parameter in Type~Ia Supernovae

### 2016-09-07 forgerard.tex
respond to Greg comments

\title{Evidence for an Intrinsic Color Parameter in Type~Ia Supernovae}
\author
{
    A.~G.~Kim,\altaffilmark{1}
    G.~Smadja,\altaffilmark{2} 
    G.~Aldering,\altaffilmark{1}
    P.~Antilogus,\altaffilmark{4}
    S.~Bailey,\altaffilmark{1}
    C.~Baltay,\altaffilmark{5}
    K.~Barbary,\altaffilmark{3}
    D.~Baugh,\altaffilmark{6}
    K.~Boone,\altaffilmark{1,3}
    S.~Bongard,\altaffilmark{4}
    C.~Buton,\altaffilmark{2}
    J.~Chen,\altaffilmark{6}
    N.~Chotard,\altaffilmark{2}
    Y.~Copin,\altaffilmark{2}
    P.~Fagrelius,\altaffilmark{1,3}
    H.~K.~Fakhouri,\altaffilmark{1,3}
    U.~Feindt,\altaffilmark{9}
    D.~Fouchez,\altaffilmark{10}
    E.~Gangler,\altaffilmark{11}  
    B.~Hayden,\altaffilmark{1}
    W.~Hillebrandt,\altaffilmark{16}
    M.~Kowalski,\altaffilmark{9,12}
    P.-F.~Leget,\altaffilmark{11}
    S.~Lombardo,\altaffilmark{9}
    J.~Nordin,\altaffilmark{1,9}
    R.~Pain,\altaffilmark{4} 
    E.~Pecontal,\altaffilmark{13}
    R.~Pereira,\altaffilmark{2}
    S.~Perlmutter,\altaffilmark{1,3}
    D.~Rabinowitz,\altaffilmark{5}
    M.~Rigault,\altaffilmark{9} 
    D.~Rubin,\altaffilmark{1,14}
    K.~Runge,\altaffilmark{1}
    C.~Saunders,\altaffilmark{1,3}
   R.~Scalzo,\altaffilmark{8}
    C.~Sofiatti,\altaffilmark{1,3} 
    N.~Suzuki,\altaffilmark{1}
    S.~Taubenberger,\altaffilmark{16}
    C.~Tao,\altaffilmark{6,10}
    R.~C.~Thomas,\altaffilmark{1} \\
    (The Nearby Supernova Factory)
}

@@

\altaffiltext{8}
{
    Research School of Astronomy and Astrophysics,
    Australian National University,
    Canberra, ACT 2611, Australia
}
\altaffiltext{9}
{
    Institut fur Physik,  Humboldt-Universitat zu Berlin,
    Newtonstr. 15, 12489 Berlin
}
\altaffiltext{10}
{
    Centre de Physique des Particules de Marseille, 
    Aix-Marseille Universit\'e , CNRS/IN2P3, 
    163 avenue de Luminy - Case 902 - 13288 Marseille Cedex 09, France
}
\altaffiltext{11}
{
    Clermont Universit\'e, Universit\'e Blaise Pascal, CNRS/IN2P3, Laboratoire de Physique Corpusculaire,
    BP 10448, F-63000 Clermont-Ferrand, France
}
\altaffiltext{12}
{
    DESY, D-15735 Zeuthen, Germany
}
\altaffiltext{13}
{
    Centre de Recherche Astronomique de Lyon, Universit\'e Lyon 1,
    9 Avenue Charles Andr\'e, 69561 Saint Genis Laval Cedex, France
}
\altaffiltext{14}
{
    Department of Physics, Florida State University,
    315 Keen Building, Tallahassee, FL 32306-4350
}
\altaffiltext{15}
{
    Computational Cosmology Center, Computational Research Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road MS 50B-4206, Berkeley, CA, 94720
}

@@

\keywords{supernovae: general}

@@

\multicolumn{3}{c}{} &  \colhead{(km s$^{-1}$)} & \multicolumn{5}{c}{}

@@

The best-fit distributions for $R_{\delta X}=\frac{\delta_X}{\delta_1-\delta_2}$ are shown in Figure~\ref{rx:fig}.
Although the posteriors of $\delta$ are close to Gaussian, $\delta_1-\delta_2$ can approach zero, resulting in the
elongated tail in negative  $R_{\delta X}$.
All values are negative and inconsistent with zero: a color excess in $B-V$ produces a brightening of the supernova.
The values of $\delta_X/\delta_V-1$ are not monotonic with color: they increase from the $U$ to $V$, and then decrease
from $V$ to $I$,
a consequence
of which is that
a reddening in $B-V$ comes with a blueing in $V-R$.
These are not the behaviors expected from normal dust,
circumstellar dust suggested as a source of SN extinction \citep{2005ApJ...635L..33W,2008ApJ...686L.103G}, or a blackbody in the Raleigh-Jeans tail, and so may be indicative of
more complicated underlying supernova physics.

@@

We plot $E_{eff}(B-V) + const$ versus
$A_{B, eff}$ for the SNf sample in Figure~\ref{rveff:fig}.
Overplotted is the expected extinction due to a pure extrinsic extinction with slope equal to $\gamma_1$.
The constant added to the effective color excess
corresponds to specifying a specific color as having zero extinction, and is chosen  qualitatively to form a ridge line 
bounding the high-extinction side of the data; the discussion that follows does not depend on the specific choice of the constant.
The right plot of Figure~\ref{rveff:fig} shows  how much more extinction (relatively) is
observed compared to the expected extinction from extrinsic dust for the corresponding effective color excess,
 $\Delta A_B =  A_{B,eff}- \gamma_1 E_{eff}(B-V)$.
Over all effective color excesses the difference is capped by the ridge line, except for extremely
blue supernovae that lie slightly but noticeably above the ridge line.
However, with increasing effective color excess there is a broadening of the distribution toward lower observed extinctions.
A linear fit between $E_{eff}(B-V)$  and $\Delta A_B$ gives a slope of $-0.33$.

@@

@@

SN~2014J   is one of several SNe~Ia that exhibits colors that imply a low $R_V<2.0$ \citep{2014ApJ...788L..21A, 2014MNRAS.443.2887F, 2015ApJ...805...74B}.
The colors of this supernova can be studied in comparison to SN~2011fe, an object with similar
spectral evolution that 
suffers low Galactic and host reddening
\citep[e.g.][]{2006MNRAS.369.1880E,2007AJ....133...58K,2008MNRAS.384..107E,2010AJ....139..120F, 2014ApJ...788L..21A}.
The overall higher photospheric velocities of
SN~2014J are unimportant  based on the findings of  \S\ref{results:sec}.

@@

\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F_V<3$ encountered by the SN~Ia population.
\citet{2011ApJ...729...55F} and the above authors
find that supernovae with large $E_o(B-V)$ 
preferentially exhibit low $R^F_V$. 
Our model predicts both of these effects with a single form of extrinsic dust, as is borne out by the
SNf sample shown in Figure~\ref{rveff:fig}. 
Our model also predicts that the converse also holds: supernovae with the smallest $E_o(B-V)$ due to intrinsic color
will appear overly faint when analyzed within the context of a canonical dust.
The trend in Figure~\ref{rveff:fig} showing an increased range of brighter-than-expected magnitudes  with increasing $E_{eff}(B-V)$  can
be compared
to Figure~\ref{betoule:fig}, which shows the $B$-band  Hubble residuals as a function of the color $C$ parameter (directly
comparable to $E(B-V)$) from the Hubble diagram analysis of \citet{2014A&A...568A..22B}.
A linear fit for the relationship between color and Hubble residual has a significant slope ($-0.67$), as expected.

>>>>

\title{Evidence for an Intrinsic Color Parameter in Type~Ia Supernovae}
\author
{
    A.~G.~Kim,\altaffilmark{1}
    G.~Smadja,\altaffilmark{2} 
    G.~Aldering,\altaffilmark{1}
    P.~Antilogus,\altaffilmark{4}
    S.~Bailey,\altaffilmark{1}
    C.~Baltay,\altaffilmark{5}
    K.~Barbary,\altaffilmark{3}
    D.~Baugh,\altaffilmark{6}
    K.~Boone,\altaffilmark{1,3}
    S.~Bongard,\altaffilmark{4}
    C.~Buton,\altaffilmark{2}
    J.~Chen,\altaffilmark{6}
    N.~Chotard,\altaffilmark{2}
    Y.~Copin,\altaffilmark{2}
    P.~Fagrelius,\altaffilmark{1,3}
    H.~K.~Fakhouri,\altaffilmark{1,3}
    U.~Feindt,\altaffilmark{9}
    D.~Fouchez,\altaffilmark{10}
    E.~Gangler,\altaffilmark{11}  
    B.~Hayden,\altaffilmark{1}
    W.~Hillebrandt,\altaffilmark{16}
    M.~Kowalski,\altaffilmark{9,12}
    P.-F.~Leget,\altaffilmark{11}
    S.~Lombardo,\altaffilmark{9}
    J.~Nordin,\altaffilmark{1,9}
    R.~Pain,\altaffilmark{4} 
    E.~Pecontal,\altaffilmark{13}
    R.~Pereira,\altaffilmark{2}
    S.~Perlmutter,\altaffilmark{1,3}
    D.~Rabinowitz,\altaffilmark{5}
    M.~Rigault,\altaffilmark{9} 
    D.~Rubin,\altaffilmark{1,14}
    K.~Runge,\altaffilmark{1}
    C.~Saunders,\altaffilmark{1,3}
    C.~Sofiatti,\altaffilmark{1,3} 
    N.~Suzuki,\altaffilmark{1}
    S.~Taubenberger,\altaffilmark{16}
    C.~Tao,\altaffilmark{6,10}
    R.~C.~Thomas,\altaffilmark{1} \\
    (The Nearby Supernova Factory)
}

@@

\altaffiltext{9}
{
    Institut fur Physik,  Humboldt-Universitat zu Berlin,
    Newtonstr. 15, 12489 Berlin
}
\altaffiltext{10}
{
    Centre de Physique des Particules de Marseille, 
    Aix-Marseille Universit\'e , CNRS/IN2P3, 
    163 avenue de Luminy - Case 902 - 13288 Marseille Cedex 09, France
}
\altaffiltext{11}
{
    Clermont Universit\'e, Universit\'e Blaise Pascal, CNRS/IN2P3, Laboratoire de Physique Corpusculaire,
    BP 10448, F-63000 Clermont-Ferrand, France
}
\altaffiltext{12}
{
    DESY, D-15735 Zeuthen, Germany
}
\altaffiltext{13}
{
    Centre de Recherche Astronomique de Lyon, Universit\'e Lyon 1,
    9 Avenue Charles Andr\'e, 69561 Saint Genis Laval Cedex, France
}
\altaffiltext{14}
{
    Department of Physics, Florida State University,
    315 Keen Building, Tallahassee, FL 32306-4350
}
\altaffiltext{15}
{
    Computational Cosmology Center, Computational Research Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road MS 50B-4206, Berkeley, CA, 94720
}

@@

\keywords{supernovae: genera; supernovae:SN 2014J}

@@

 & (\AA) & (\AA)&  \colhead{(km s$^{-1}$)} & \multicolumn{5}{c}{}

@@

The best-fit distributions for $R_{\delta X}=\frac{\delta_X}{\delta_1-\delta_2}$ are shown in Figure~\ref{rx:fig}.
Although the posteriors of $\delta$ are close to Gaussian, $\delta_1-\delta_2$ can approach zero, resulting in the
elongated tail in negative  $R_{\delta X}$.
All values are negative and inconsistent with zero: a color excess in $B-V$ produces a brightening of the supernova.
The values of $\delta_X/\delta_V-1$ are not monotonic with color: they increase from the $U$ to $V$, and then decrease
from $V$ to $I$,
a consequence
of which is that
a reddening in $B-V$ comes with a blueing in $V-R$.
These are not the behaviors expected from normal dust,
circumstellar dust suggested as a source of SN extinction \citep{2005ApJ...635L..33W,2008ApJ...686L.103G,
2015ApJ...807L..26G}, or a blackbody in the Raleigh-Jeans tail, and so may be indicative of
more complicated underlying supernova physics.

@@

We plot $E_{eff}(B-V) + const$ versus
$A_{B, eff}$ for the SNf sample in Figure~\ref{rveff:fig}.
Overplotted is the expected extinction due to a pure extrinsic extinction with slope equal to $\gamma_1$.
The constant added to the effective color excess
corresponds to specifying a specific color as having zero extinction, and is chosen  qualitatively to form a ridge line 
bounding the high-extinction side of the data; the discussion that follows does not depend on the specific choice of the constant.
The right plot of Figure~\ref{rveff:fig} shows  how much more extinction (relatively) is
observed compared to the expected extinction from extrinsic dust for the corresponding effective color excess,
 $\Delta A_B =  A_{B,eff}- \gamma_1 E_{eff}(B-V)$.
Over all effective color excesses the difference is capped by the ridge line, except for extremely
blue supernovae that lie slightly but noticeably above the ridge line.
However, with increasing effective color excess there is a broadening of the distribution toward lower observed extinctions.
A linear fit between $E_{eff}(B-V)$  and $\Delta A_B$ gives a slope of $-0.45 \pm 0.11$.

@@

The range of SiII equivalent widths is $\pm 20$~\AA\ whereas the width of the $B$-band is 851~\AA.  
The implied span in $B$ magnitude based on $\beta_1$ is 0.54~mag.  Therefore $\beta_1$ cannot wholly be attributed to the line itself.
Similarly, the CaII equivalent widths have range $\pm 50$~\AA, while the width of the $U$ band is
701~\AA.  The implied span in $U$ magnitude is is 0.21~mag, so $\alpha_0$ cannot be completely due to the line iteslf.

@@

SN~2014J   is one of several SNe~Ia that exhibits colors that imply a low $R_V<2.0$ \citep{2014ApJ...788L..21A, 2014MNRAS.443.2887F, 
2014ApJ...795L...4K, 2015ApJ...805...74B}.
The colors of this supernova can be studied in comparison to SN~2011fe, an object with similar
spectral evolution that 
suffers low Galactic and host reddening
\citep[e.g.][]{2006MNRAS.369.1880E,2007AJ....133...58K,2008MNRAS.384..107E,2010AJ....139..120F, 2014ApJ...788L..21A}.
The overall higher photospheric velocities of
SN~2014J are unimportant  based on the findings of  \S\ref{results:sec}.

@@

\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F_V<3$ encountered by the SN~Ia population.
\citet{2011ApJ...729...55F} and the above authors
find that supernovae with large $E_o(B-V)$ 
preferentially exhibit low $R^F_V$. 
Our model predicts both of these effects with a single form of extrinsic dust, as is borne out by the
SNf sample shown in Figure~\ref{rveff:fig}. 
Our model also predicts that the converse also holds: supernovae with the smallest $E_o(B-V)$ due to intrinsic color
will appear overly faint when analyzed within the context of a canonical dust.
The trend in Figure~\ref{rveff:fig} showing an increased range of brighter-than-expected magnitudes  with increasing $E_{eff}(B-V)$  can
be compared
to Figure~\ref{betoule:fig}, which shows the $B$-band  Hubble residuals as a function of the color $C$ parameter (directly
comparable to $E(B-V)$) from the Hubble diagram analysis of \citet{2014A&A...568A..22B}.
A linear fit for the relationship between color and Hubble residual has a significant slope ($-0.57 \pm 0.07$), consistent with the slope
found for the SNfactory data.
